\cleardoublepage
% \newpage
% \thispagestyle{empty}
% \mbox{}

\chapter*{Introduction}
\label{ch:chapter1}

\section*{Motivation}


In recent years, energy consumption has been revealed as one of the
barriers that has stopped performance improvements in all computer systems,
from mobile devices to big data centres. In order to improve the energy
efficiency, it is necessary to exploit greater specialisation levels in the
hardware. Nowadays, architectures provide different types of heterogeneity,
from heterogeneous configurations for data centres, to processors with
hardware accelerators inside the chip, hybrid platforms composed by general
purpose processors (CPUs) and specific purpose accelerators (like GPUs,
Intel Xeon Phi, etc.), asymmetric multicores (AMPSs) with the same
instruction set, or heterogeneous multicores with different instruction
sets~\cite{NRC2011,KoSh13}.

% En los últimos años, el consumo energético ha emergido como una de las
% principales barreras que ha frenado la mejora de rendimiento en los
% sistemas de cómputo, desde dispositivos móviles hasta grandes centros de
% procesamiento de datos (CPDs). Con el fin de mejorar la eficiencia
% energética, resulta necesario explotar mayores grados de especialización en
% el hardware. Las arquitecturas actuales proporcionan diversos tipos de
% heterogeneidad, que van desde configuraciones heterogéneas en CPDs, hasta
% procesadores con aceleradores hardware integrados en el propio chip,
% plataformas híbridas formadas por procesadores de propósito general (CPUs)
% y aceleradores de propósito específico (como pueden ser GPUs, Intel Xeon
% Phi, etc.), multicores asimétricos (AMPSs) con repertorio común de
% instrucciones o multicores heterogéneos con múltiples repertorios de
% instrucciones~\cite{NRC2011,KoSh13}.


The correct exploitation of performance and energy efficiency offered by
these architectures leads to an increase in complexity of the software that
permits, in the most transparent way for the end user, the adaptation of
already existing applications or newly developed applications to increase
performance and reduce the energy consumption~\cite{OsTo10, SKC+15}. One of
the main challenges in this topic is the right assignment of work to the
existent computational resources, which should be done in such a way that,
simultaneously, its activity degree and the adaptation between the task and
the resources are optimised. In response to these needs, one of the
scientific community solutions proposed is the development of programming
models which expose the task parallelism, exploited in execution
time by task schedulers (usually known as runtimes~\cite{VMC+14}).  This
software layer can manage, in a transparent user way, the process of
managing the dependencies between tasks, data transfers between
heterogeneous environments, and the efficient assignment between types of
tasks and processors, as many others aspects.

% La correcta explotación del rendimiento y la eficiencia energética
% ofrecidos por estas arquitecturas conlleva, de forma necesaria, un
% incremento en la complejidad del software que permita, de forma lo más
% transparente posible para el usuario final, la adaptación de aplicaciones
% ya existentes o de nuevo desarrollo para aumentar el rendimiento y
% reducir su consumo energético~\cite{OsTo10,SKC+15}.Uno de los principales
% retos en este sentido es la correcta asignación de trabajo a los recursos
% computacionales existentes, de modo que, de forma simultánea, se optimice
% su grado de ocupación y la adaptación del recurso escogido a la tarea
% específica a desarrollar.En respuesta a estas necesidades, una de las
% soluciones propuestas por la comunidad científica es el desarrollo de
% modelos de programación que expongan el paralelismo a nivel de tareas,
% explotado en tiempo de ejecución por planificadores de tareas (comúnmente
% conocidos como runtimes~\cite{VMC+14}). Esta capa de software es capaz
% de gestionar de forma transparente al usuario los procesos de gestión de
% dependencias de datos entre tareas, las transferencias de datos en
% entornos heterogéneos, y el mapeado eficiente entre tipos de tareas y de
% procesador, entre otros aspectos.

Regarding underlying architectures, the work focuses on one of the family
architectures with more future perspective between the previously cited:
asymmetric multicore architectures (AMPs) with low energy consumption and,
specifically, processors of ARM's big.LITTLE family. This kind of
configurations has caused a high interest in the application market and
mobile ecosystems, because of the specialisation of each kind of processor
to specific tasks. This kind of specialisation is translated into the
activation/deactivation of each kind of core depending of the computational
demand or application priorities in execution, or the energy limitations of
each execution situation. 

% Atendiendo a las arquitecturas subyacentes, el trabajo se centra en una de
% las familias de arquitecturas con más perspectiva de futuro de entre las
% anteriormente mencionadas: procesadores multinúcleo asimétricos (AMPs) de
% bajo consumo, y en concreto, procesadores de la familia big.LITTLE de ARM.
% Este tipo de configuraciones ha suscitado gran interés en el mercado de las
% aplicaciones y ecosistemas móviles, debido a la especialización de cada
% tipo de procesador ante tareas concretas. Esta especialización se suele
% traducir en la activación/desactivación de cada tipo de núcleo en función
% de la demanda computacional o prioridad de las aplicaciones en ejecución, o
% de las limitaciones energéticas ligadas a cada escenario concreto de
% ejecución.


\section*{Objectives}
From a high performance computing (HPC) point of view, using all
computational resources together from one multithreaded application has a
special interest. Although some techniques developed on existing libraries
and task schedulers can be used in this kind of architectures, these expose
new challenges to face. This work focus on two of these challenges:

% Desde el punto de vista de la computación de altas prestaciones (HPC), el
% uso conjunto de todos los recursos computacionales desde una misma
% aplicación multihebra resulta de especial interés. Aunque algunas de las
% técnicas ya desarrolladas en bibliotecas o planificadores de tareas
% existentes pueden ser aplicadas directamente en este tipo de arquitecturas,
% éstas exponen nuevos retos y problemáticas que hay que abordar. Este
% trabajo se centra en dos de estos retos:


\begin{itemize}
\item Integration and combination of specific asymmetry-aware libraries
  with conventional \emph{runtimes}  (non conscious of the asymmetry), and
  comparative of this approach with a \emph{runtime} asymmetry-aware
  design.
% \begin{itemize}
% \item Integración y combinación de bibliotecas específicas conscientes de
%   la asimetría con \emph{runtimes} convencionales (no conscientes de la
%   asimetría), y comparativa de este enfoque con un diseño de \emph{runtime}
%   sí consciente con la asimetría de la arquitectura.

\item Developing of new energy consumption reduction techniques. Applying
  dynamic frequency scaling (DVFS) techniques, or deactivating cores in
  different moments during the application execution depending on the
  computing demands and parallelism degree available, and evaluation of the
  impact of these techniques in run-time and global energy consumption.
% \item Diseño de técnicas de reducción de consumo energético. Aplicación de
%   técnicas de escalado dinámico de frecuencia (DVFS) o desactivación de
%   núcleos en distintas fases de la ejecución de una determinada aplicación,
%   en función de las demandas de cómputo y grado de paralelismo disponible, y
%   evaluación del impacto de este tipo de técnicas en el tiempo de ejecución
%   y consumo energético global.
% \end{itemize}
\end{itemize}









\section*{Methodology and work plan}

In order to achieve the objectives described in the previous section, the
next specific tasks have been done:

% Para alcanzar los objetivos descritos en el apartado anterior, se han
% abordado las siguientes tareas específicas:
\begin{enumerate}[T1.]

\item Study of the state of the art of asymmetric systems from an
  architectural point of view, and of the software support for exploiting
  its characteristics, from the operative system level, \emph{runtime}
  level or application level.
% \item Estudio del estado del arte en sistemas asimétricos desde el punto de
%   vista arquitectural, y del soporte software disponible para explotar sus
%   características, a nivel de sistema operativo, \emph{runtime} o
%   aplicación.

\item Study, analysis and evaluation of performance of families of
  libraries and schedulers non conscious of the architecture's asymmetry, and
  also of solutions adapted for this kind of architectures. In the linear
  algebra scope, used as the main thread of this work, the BLIS
  library~\cite{BLIS1} has been used.
% \item Estudio, análisis y evaluación del rendimiento de familias de
%   bibliotecas y planificadores de tareas en tiempo de ejecución no
%   conscientes de la asimetría de la arquitectura, así como soluciones ya
%   adaptadas a esta familia de arquitecturas. En el ámbito del álgebra
%   lineal, utilizado como hilo conductor del trabajo, se ha utilizado la
%   biblioteca BLIS~\cite{BLIS1}.

\item Integration of the previous libraries over existing asymmetry-aware
  non conscious schedulers, to exploit the asymmetry of the architecture
  without modifying the actual schedulers in a transparent way, specially
  over a set of mathematical interesting operations.
% \item Integración de las anteriores bibliotecas sobre \emph{runtimes} no
%   conscientes de la asimetría ya existentes, para explotar de forma
%   transparente la asimetría de las arquitecturas sin modificaciones en los
%   planificadores, para un conjunto de operaciones matemáticas de interés.

\item Design of heuristics, resource scheduling policies and exploitation
  of low power consuming modes provided by this kind of architectures,
  which allow an optimal use of the resources in existing schedulers, with
  the objective of exploiting every existing computational resource in an
  energy efficient way.
  
% \item Diseño de heurísticas, políticas de asignación de recursos y
%   explotación de los modos de bajo consumo proporcionados por las
%   arquitecturas, que permitan un aprovechamiento óptimo de los recursos a
%   nivel del planificador de tareas en \emph{runtimes} existentes, con el
%   objetivo de explotar de manera energéticamente eficiente todos los
%   recursos computacionales existentes.


\item Evaluation of performance and energy efficiency of the new policies
  developed. Comparison between these results with current asymmetry-aware
  solutions. 
% \item Evaluación del rendimiento y eficiencia energética de las nuevas
%   políticas desarrolladas. Comparación de los resultados con soluciones
%   actuales  conscientes de la asimetría de la arquitectura.
\end{enumerate}


\section*{Document structure}
This document is organised as follow: Chapter~\ref{ch:chapter2} shows a
general description of the heterogeneous and asymmetric architectures most
used nowadays, focusing on ARM's big.LITTLE architectures used in this
work. Also, specific characteristics of the environment used for testing
(including hardware and software) are described. Chapter~\ref{ch:chapter3}
introduces the two parallel strategies used in this work: task-level
parallelism and data-level parallelism, focusing on the characteristics of
each one of them and also on their advantages and disadvantages.
Chapters~\ref{ch:chapter4}~and~\ref{ch:chapter5} show a complete
description of the work developed, splitted into the proposed techniques
and their evaluation for obtaining a performance gain over task-level
schedulers, and techniques developed for obtain at least some energy
efficiency gain, respectively. Finally, chapter~\ref{ch:chapter6} shows a
set of general conclusions and future work.


% La presente memoria se estructura como sigue. 
% El Capítulo~\ref{ch:chapter2} ofrece una descripción general de las
% arquitecturas heterogéneas y asimétricas actualmente más utilizadas en el
% ámbito de la computación de altas prestaciones, haciendo especial hincapié
% en el paradigma big.LITTLE de ARM utilizado en el trabajo
% realizado. Además, detalla las características específicas del entorno
% experimental empleado a nivel hardware y software.
% El Capítulo~\ref{ch:chapter3} introduce dos de las estrategias de
% paralelización utilizadas en el desarrollo del trabajo: extracción de
% paralelismo a nivel de tareas, y extracción de paralelismo a nivel de
% datos, centrándose en las características de cada uno, así como en sus
% ventajas e inconvenientes.
% Los Capítulos~\ref{ch:chapter4} y~\ref{ch:chapter5} describen en detalle el
% trabajo realizado, dividido en la propuesta y evaluación de técnicas para
% la mejora de rendimiento en planificadores a nivel de tareas, y en el
% desarrollo de soluciones para la mejora de la eficiencia energética,
% respectivamente.
% Finalmente, el Capítulo~\ref{ch:chapter6} incluye una serie de conclusiones
% generales y líneas de trabajo futuras.


%-- Configuraciones para emacs --
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "./principal.tex"
%%% End:
